\documentclass[12pt]{article}

\title{College Flex Day }
\author{Henry Oehlrich}

\usepackage{booktabs}
\usepackage[margin=1in]{geometry}

\begin{document}
\maketitle
\subsection*{What is the universal language of the world.}


\subsection*{If you were a plant, what plant would you be? Why?}
Elm tree.

\subsection*{If you were an ice cream, what ice cream would you be? What do you want to be? How do you know?}

\subsection*{Are you a circle, a square, a triangle, or a squiggle?}
I am a triangle.

\subsection*{List concrete items you spend a lot of time with and abstract items you feel people don't define correctly.}

\begin{tabular}{l|l}
    \toprule
    Concrete & Abstract \\
    \midrule
    Laptop & Loud \\
    Pencil and eraser & Programmer \\
\end{tabular}

\subsection*{What does it feel like to do your favorite activity?} 
I'm neither a talented artist nor an experienced musician, and I'm certainly
not a good singer, but when I'm in the zone, behind my computer monitor,
coding, my hands never leaving the keyboard, I feel like a wizard, coconut oily
smooth, gliding from file to file, refactoring on the fly. When I slip into the
zone like that I can program for hours straight without food, sleep, or human
contact. At 6pm I start working on a simulation and pathfinding program for my
school's Science Olympiad robotics event. The next thing I know, it's 1am and
the simulation is running smoothly and paths are generating properly.

\subsection*{What do you loathe and love about yourself}
I love how I can grasp a concept quickly and understand it fully. I loathe how
I sometimes explain things to people. I've been told a few times that I can be
a little condescending.

\end{document}

