\documentclass{article}

\title{APUSH Final Paper}
\author{Henry Oehlrich}

\begin{document}
\maketitle{}

The stock market crash in October of 1929 and the following depression of the
1930’s are often used as examples of failures of capitalism and unregulated
markets. It is not proven that the stock market crash mandated the mass
unemployment that followed. It instead could be argued that the government’s
actions in the 20’s and response to the downturn in the 30’s were what caused
the catastrophe that was the Great Depression. The Great Depression was
primarily caused by the unsustainable credit expansion and subsequent boom-bust
cycle fueled by the Federal Reserve. Artificially low interest rates of the
1920s led to excessive credit creation, bad investments, and misallocations of
resources, creating an unsustainable economic boom. The contractionary
monetary policies pursued by central banks after the stock market crash in 1929
further deepened the economic downturn, preventing necessary market adjustments
from occurring and stopping the natural recovery process.

\section{Easy money of the late 1920's}

The 1920s, often referred to as the Roaring Twenties, contained a period of
apparent prosperity and economic growth in the United States. Unfortunately,
this prosperity and booming economic activity was largely driven by
unsustainable factors. The artificial boom of the 1920s was caused by
government policies, particularly monetary interventions.

One of the main causes of the boom of the 1920s was the expansionary monetary
policies of Federal Reserve. The Fed’s decision to keep interest rates
artificially low led to an increase in bank credit and the availability of
money. For example, from 1921 to 1929, the money supply in the United States
expanded by 60\%. This influx of easy credit, nicknamed ”Easy money”,
stimulated economic activity and contributed to the appearance of a booming
economy.

The availability of loans and the optimistic outlook of investors, due to a
false boom, fueled speculative excesses, particularly in the stock market. This
resulted in a rapid increase in stock prices and trading volumes, with the Dow
Jones Industrial Average reaching unprecedented levels. For example, between
1924 and 1929, the Dow rose by over 200\% This overspeculation created a bubble
and set the stage for a massive stock market crash.

The artificial boom of the 1920s was further propelled by excessive consumer
spending and reliance on debt. Readily available loans and installment buying
(going into debt to purchase something) allowed individuals to engage in higher
levels of consumption. Because of the volume of credit, much of this spending
was financed by credit rather than real savings. For example, consumer debt
rose from \$1.8 billion in 1920 to \$7 billion in 1929.4 This unsustainable
consumer debt burden left people and assets vulnerable when the economic
downturn hit.

\section{Interventionist response policies}

In October of 1929, the stock market crashed. Although unemployment rose
immediately after the crash, the unemployment rate peaked at 9\% two months
after the crash, and then began to trend downward. By June of 1930, 8 months
later, unemployment was down to 6\%. By recession standards for the time, this
was not extraordinary. In 1920 there was a similar downturn and unemployment
hit 11\% in 8 months before recovering quickly. The difference between these
two downturns was government intervention. President Woodrow Wilson and
Warren Harding shared the depression of the early 20s. Both Wilson and Harding
did nothing to remedy the economy and it recovered quickly. On the other hand
Presidents Hoover and Roosevelt engaged in large amounts of government spending
in an attempt to stimulate the economy. After a series of unprecedented
government interventions, unemployment rate shot to over 20\% for 3 years
straight.

In early June of 1930, unemployment was at 6\% and trending downward. The
government, feeling compelled to do something, passed into law the Smoot-Hawley
Tariffs. The tariffs were designed to reduce unemployment by restricting
imports to the United States so that more goods were produced domestically by
American workers. Although well meaning, it had disastrous effects. 1008
economists signed a public appeal to not pass the bill stating that the bill
would not reduce unemployment [citation needed]. Instead, the economists
argued, it would lead to retaliation that would make it harder for Americans to
sell their goods in other countries. As predicted, by imposing large tariffs,
the Act sparked retaliation from other nations, leading to a decline in
international trade. Dartmouth economist Douglas Irwing wrote that the tariff
has ”become synonymous with an avalanche of protectionism that led to the
collapse of world trade and the Great Depression” [citation needed]. In the
following four months after passing the tariffs, unemployment rose to 14\%.
Unemployment stayed in the double digits until the end of the depression in
1939.

Herbert Hoover was president at the start of the Great Depression. Hoover,
having previously been the Secretary of Commerce, believed that the stock
market crash and the downturn as a whole was caused by overspeculation of
investors buying on credit. In an attempt to prevent panic from spreading
throughout the economy, he told business leaders to maintain wages and
collected money from private business in order to stimulate the economy. He
believed that financial losses should affect profits, not employment. His
response was nearly entirely based on spending copious amounts of money in
order to stimulate the economy. Government and business had spent more in the
first six months of 1930 than in the entire previous year. [citation needed]
Although the efficacy of Hoover’s numerous remedies was limited, Hoover’s
defenders praise his attempts to fix the economy. The Herbert Hoover
Presidential Library and Museum defends Hoover stating that ”no one in his
place could have done more” and that ”very few of his predecessors have done as
much.”

\section{Failure of the Federal Reserve System}

The Federal Reserve was established in 1913 in response to financial panics in
the late 19th and early 20th century. Its duty was to act as a lender of last
resort during times of financial stress. It would (theoretically) accomplish
this by providing liquidity, or cash, to banks. The Fed was designed to
stabilize the financial system and prevent widespread bank runs.

In their book \textit{A Monetary History of the United States}, Milton Friedman
and Anna Schwartz propose an explanation for the Great Depression. They
determine that the Depression was ultimately caused by a deflation of the money
supply from the failure of the Federal Reserve to save the banking system. From
1929 to 1933, one third of the nations banks (around 7,500) closed due to
failure or merger. Resulting from this, the nations money supply, defined as
bank deposits combined with currency circulation, shrank by a third. The irony,
Friedman and Schwartz argue, is that the Federal Reserve was created to serve
the very purpose it declined to serve.

In the early years of the Great Depression, the Federal Reserve enacted policy
that supported monetary contraction. Instead of providing sufficient money to
the banking system, the Fed restricted the money supply and raised interest
rates. These actions further contracted the economy and exacerbated the banking
crisis. By not acting as a lender of last resort, the Federal Reserve failed to
provide the necessary money to stabilize the financial system and prevent
widespread bank failures.

As the depression progressed, there were numerous bank runs. Panicked depositors,
now unsure of the safety of their money, rushed to withdraw their money from banks.
These runs caused banks to experience severe shortages, leading to bank failures. The
Federal Reserve, as the lender of last resort, had the authority, ability, and duty to
provide emergency loans to troubled banks to prevent the crisis. However, the Fed’s
response was inadequate; it did not provide sufficient money to overrun banks in need.

\end{document}
