\documentclass{article}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{caption}

\captionsetup[figure]{labelfont=bf}

\author{Henry Oehlirch \and Lahari Jaligama}
\title{Difference between mean typing speeds of STEM and humanities students}

\begin{document}
\maketitle{}

\section{Question, Hypothesis, \& Data Schema}

Is there a statistically significant difference between the mean typing speed
for students who consider themselves more STEM-oriented and those who consider
themselves more humanities-oriented?

\begin{gather}
    H_0: \overline{STEM} = \overline{Humanities} \\
    H_a: \overline{STEM} \neq \overline{Humanities}
\end{gather} 

\begin{figure}[h]
    \centering
    \begin{tabular}{lllll}
        \toprule
        i & Field & Test 1 & Test 2 & Test 3 \\
        \midrule
        1 & $field_1$ & $speed1_1$ & $speed2_1$ & $speed3_1$ \\
        ... & ... & ... & ... & ... \\
        1 & $field_n$ & $speed1_n$ & $speed2_n$ & $speed3_n$ \\
    \end{tabular}
    \caption{Data schema}
\end{figure}

\section{Data Collection \& Analysis}

We will collect data from a random sample of seniors at Raleigh Charter. A
survey will be sent out to 50 seniors, asking them to take three predetermined,
30-word typing tests and report their words per minute for each test. The
survey will also ask for the student to self-report their field of interest:
STEM or humanities.

We predict a sampling bias in the data, as students who are more confident in
their typing speed may be more likely to respond to the survey. It is
reasonable to assume that this bias will be present in both groups, so we do
not expect it to affect the results of the analysis.

The survey will be sent out during the earliest stats class that doesn't
coincide with a major senior AP test (so hopefully Thursday).

\subsection{Questions}

This survey is anonymous. For each test below, attempt the test once and record
the speed in words per minute (WPM) in the corresponding field. \\

\noindent
\verb|Words per minute (WPM) on test one {link}: _ | \\\\
\verb|Words per minute (WPM) on test two {link}: _ | \\\\
\verb|Words per minute (WPM) on test three {link}: _ | \\\\
\verb|Field of interest: STEM or humanities? {link}: {STEM/humanities} | \\

We do not expect any bias in the survey responses. Each question is objective
and, because the form is anonymous, there is no incentive for students to
misreport their typing speed or field of interest.

\subsection{Analysis}

We will use a two-sample t-test to determine if there is evidence to conclude
that the mean typing speed of STEM students is different from the mean typing
speed of humanities students. The validity of the t-test is contingent on the
sizes of the two samples. Because there are only 120 seniors at Raleigh
Charter, it is likely that we will break the 10\% condition for the t-test.
This can only be determined after the data is collected because we do not know
the response rate or the distribution of responses between the fields.

\end{document}
