\documentclass{article}

\author{Henry Oehlrich}
\title{In Defense of Liberty}

\begin{document}

\maketitle

The State has demonized liberty. From their high horse of supposed economic
equality, social justice, and self-righteousness, the State has discredited
champions of personal freedom by calling them selfish, racist, and immoral.
This occurs because the State judges decisions based off intentions (as opposed
to results). Much of the public, who decides in part who is in power, is wholly
unaware of the results of decisions and also judges them based on intentions.
On the contrary, liberty minded individuals recognize that general prosperity
is always and only a result of creative and economic liberty.

All the most powerful nations have achieved their position due to the
innovation and productivity that competition and individual liberty due to
limited government provides. When freedom is stemmed and the State controls the
natural systems of society, fights over trivial policy are replaced by fights
over food and water. Every communist experiment in the 20th century has proved
as much. After a nation has established itself through solid, constructive,
principles, it often finds itself enamored with promises of equity and fairness
and proceeds to destroy the very system that granted it prosperity. This is the
disease of the modern State.

Despite the policy of the central planners, business continues to grow. Growth
of business is stunted and, as regulations and taxes continue to rise,
incentives for businessmen to continue their risky endeavours decreases.

\end{document}
